% Upper-case    A B C D E F G H I J K L M N O P Q R S T U V W X Y Z
% Lower-case    a b c d e f g h i j k l m n o p q r s t u v w x y z
% Digits        0 1 2 3 4 5 6 7 8 9
% Exclamation   !           Double quote "          Hash (number) #
% Dollar        $           Percent      %          Ampersand     &
% Acute accent  '           Left paren   (          Right paren   )
% Asterisk      *           Plus         +          Comma         ,
% Minus         -           Point        .          Solidus       /
% Colon         :           Semicolon    ;          Less than     <
% Equals        =           Greater than >          Question mark ?
% At            @           Left bracket [          Backslash     \
% Right bracket ]           Circumflex   ^          Underscore    _
% Grave accent  `           Left brace   {          Vertical bar  |
% Right brace   }           Tilde        ~

%
% NSF Biographical Sketch: Dr. Kathryn D. Huff
%
% Copyright (c) 2013 by Theodore P. Pavlic
% All rights reserved.
%

\documentclass[svgnames,11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage{times}
\renewcommand\thesection{(\alph{section})}
\renewcommand\thesubsection{(\alph{subsection})}

\usepackage{calc}

\usepackage{lastpage}
%\newcommand{\pageoflastpage}{Page {\thepage} of \pageref*{LastPage}}
%\newcommand{\pageoflastpage}{{\thepage} of \pageref*{LastPage}}
\newcommand{\pageoflastpage}{}

% Redefine the plain pagestyle for the title page
\makeatletter
\let\oldps@plain\ps@plain
\renewcommand{\ps@plain}{\oldps@plain%
\renewcommand{\@evenfoot}{\hfil\pageoflastpage\hfil}%
\renewcommand{\@oddfoot}{\@evenfoot}}
\makeatother

% Use fancy for non-title pages
\usepackage{fancyhdr}
\fancyhead{}
\fancyfoot{}
\cfoot{\pageoflastpage}
\pagestyle{fancy}

\usepackage{xspace}

\usepackage[shortlabels]{enumitem}
\usepackage{graphicx}
\usepackage{varioref}
\usepackage[%
        colorlinks=true,urlcolor=,
        pdfpagelabels=true,hypertexnames=true,
        plainpages=false,naturalnames=true,
        ]{hyperref}
%
\usepackage{url}
\newcommand\doilink[1]{\href{http://dx.doi.org/#1}{#1}}
\newcommand\doi[1]{doi:\doilink{#1}}

\makeatletter
\newlength{\bibhang}
\setlength{\bibhang}{1em}
\newlength{\bibsep}
 {\@listi \global\bibsep\itemsep \global\advance\bibsep by\parsep}
\newlist{bibsection}{itemize}{3}
\setlist[bibsection]{label=,leftmargin={\bibhang+\widthof{[9]}},%
        itemindent=-\bibhang,
        itemsep=\bibsep,parsep=\z@,partopsep=0pt,
        topsep=0pt}
\newlist{bibenum}{enumerate}{3}
\setlist[bibenum]{label=\textbf{\arabic*.},leftmargin={\bibhang+\widthof{[999]}},%
        itemindent=-\bibhang,
        itemsep=\bibsep,parsep=\z@,partopsep=0pt,
        topsep=0pt}
\let\oldendbibenum\endbibenum
\def\endbibenum{\oldendbibenum\vspace{-.6\baselineskip}}
\let\oldendbibsection\endbibsection
\def\endbibsection{\oldendbibsection\vspace{-.6\baselineskip}}
\makeatother

\title{%
        \vspace{-2\baselineskip}
            \normalsize
            Biographical Sketch\\
            {\large\textbf{Dr.~Kathryn D.~Huff}}\\
            \vspace{0.5\baselineskip}
            \hrule
            \vspace{0.5\baselineskip}
            Department of Nuclear Engineering, University of California - Berkeley,
            e-mail: \href{mailto:huf@berkeley.edu}{huff@berkeley.edu},
            tel: +1-281-734-1342
        \vspace{-1.5ex}
        }
\date{}
\author{}

\hypersetup{%
        pdfsubject={NSF Biographical Sketch of Dr.~Kathryn D.~Huff},
        pdfauthor={Dr. Kathryn D.~Huff},
        pdftitle={NSF Biographical Sketch of Dr.~Kathryn D.~Huff},
        pdfkeywords={}
        }

\newcommand{\ie}{i.e.\xspace}
\newcommand{\eg}{e.g.\xspace}

\begin{document}

\maketitle
\vspace{-4\baselineskip}

\subsection{Professional Preparation}

    The University of Chicago, Chicago, IL;
        Physics;
        B.S., 2004
        \\
    The University of Wisconsin, Madison, WI;
        Nuclear Engineering;
        Ph.D., 2013
        \\
    The University of California, Berkeley, CA;
        Nuclear Engineering;
        Postdoctoral Researcher, 2013--present
        \\

\subsection{Appointments}

\begin{itemize}[label={\quad 9999--9999:},leftmargin=*,itemsep=0pt]

    \item[2013--present:]
        \textbf{Postdoctoral Scholar},
        Nuclear Science and Security Consortium,
        The University of California - Berkeley,
        Berkeley, CA

    \item[2011--2013:]
      \textbf{Laboratory Graduate Fellow},
        Used Fuel Disposition Campaign,
        Argonne National Laboratory,
        Argonne, IL

    \item[2010:]
        \textbf{Laboratory Graduate Researcher},
        Fuel Cycle Options Campaign,
        Idaho National Laboratory,
        Idaho Falls, ID

    \item[2008--2011:]
        \textbf{Graduate Research Assistant},
        Computational Nuclear Engineering Research Group,
        University of Wisconsin - Madison,
        Madison, WI

    \item[2005--2008:]
      \textbf{Undergraduate Research Assistant},
        Kavli Institute for Cosmological Physics,
        The University of Chicago,
        Chicago, IL

    \item[2006:]
      \textbf{Undergraduate Research Assistant},
        Physics Department,
        University of Chile, 
        Santiago, CHILE

    \item[2003-2004:]
      \textbf{Undergraduate Research Assistant},
        Los Alamos Neutron Science Center, LANSCE--3,
        Los Alamos National Laboratory,
        Los Alamos, NM

\end{itemize}

\subsection{Products}

\begin{bibenum}[itemsep=5pt]
   \item Aruliah,~D.A., Brown,~C.T., Chue Hong,~N.P., Davis,~M., Guy,~R.T., 
          Haddock,~S.H.D., \textbf{Huff,~K.}, Mitchell,~I., Plumbley,~M., Waugh,~B., 
          White,~E.P., Wilson,~G.V., and Wilson,~P.P.H.  ``Best Practices For 
          Scientific Computing.'' \textsl{PLOS Biology}, Vol 1, Issue 12, 2014. 
          arXiv:1210.0530 [cs.MS].

   \item \textbf{Huff K.}, Scopatz, A., Preston, N., Wilson, P.P.H. ``Rapid
          Peer Education of a Computational Nuclear Engineering Skill Suite.''
          \emph{Transactions of the American Nuclear Society Annual Conference.}
          Hollywood, FL. June 2011.
\end{bibenum}

\subsection{Synergistic Activities}

\begin{bibenum}[itemsep=4pt]

    \item \textbf{Conference service:}
        (i) Technical Program Co-Chair for the Scientific Computing with Python 
        Conference (SciPy), Austin, TX, 2013\&2014.

    \item \textbf{Community service:}
        (i) Editor, Proceedings of SciPy Scientific Computing with Python 
        Conference, 2013.
        (ii) Workshop Organizer for Software Carpentry, 2011.
        (iii) Author, panelist, The inSCIght Scientific Computing Podcast.
        (iv) Co-Founder, President, The Hacker Within Scientific Computing Group, 
        Peer-driven skill sharing. 2009-2011.

    \item \textbf{Mentoring:}
        (i) Mentoring graduate student software projects in the Reactor Design, 
        Neutronics, and Computational Methods group. 2013--present.  
        (ii) Participant, University of California -- Berkeley Py4Science Peer 
        learning group. 2013--present. 
        (iii) Curriculum Developer, Seminar Organizer, President, The Hacker 
        Within Scientific Computing Group, Peer-driven skill sharing. 2009-2011.

    \item \textbf{Teaching:}
        (i) Instructor, Curriculum Developer, Workshop Organizer for 
        Software Carpentry. 10+ workshops. 2011 -- present.
        (ii) Instructor, Curriculum Developer, Workshop Organizer for 
        The Hacker Within Scientific Computing Group. 4 workshops. 2009--2011.


\end{bibenum}

\subsection{Collaborators \& Other Affiliations}

\begin{bibsection}
    \item \textbf{Collaborators and Co-Editors:}
      \emph{The University of Wisconsin}:
            Paul P.H Wilson;
            Matthew J. Gidden;
            Anthony M. Scopatz;
            Robert W. Carlsen;
      \emph{The University of California - Berkeley}:
            Jasmina A. Vujic;
            Per F. Peterson;
            Ehud Greenspan;
            Massimiliano Fratoni;
            David L. Krumwiede;
            Charalampos Andreades;
            Anselmo T. Cisneros;
            Lakshana Huddar;
            Michael R. Laufer;
            Madicken Munk;
            Raluca O. Scarlat;
            Jeffrey E. Seifried;
            Nicolas Zweibaum;
      \emph{Lawrence Livermore National Laboratory}:
            Harris Greenberg;
      \emph{Software Carpentry}:
            Gregory V. Wilson;
    \item \textbf{Graduate Advisers and Postdoctoral Sponsors:}
        \emph{The University of Wisconsin}:
            Paul P.H. Wilson;
        \emph{Argonne National Laboratory}:
            W. Mark Nutt;
            Temitope A. Taiwo;
        \emph{University of California - Berkeley}:
            Jasmina Vujic;
            Per F. Peterson;
            Massimiliano Fratoni;
    \item \textbf{Thesis Adviser and Postgraduate-Scholar Sponsor:}
        Total graduate students: 0; Total postgraduate scholars: 0.

\end{bibsection}

\end{document}
