\documentclass[11pt]{article}
\begin{document}

\section{Publication Plans}

\begin{itemize}
\item Journal article(s) in Journal of Computers in Mathematics and Science Teaching (JCMST). JCMST is a highly respected scholarly journal that offers an in-depth forum for the interchange of information in the fields of science, mathematics, and computer science.
%
\item Conference presentations through the Special Interest Group on Computer Science Education (SIGCSE) of the Association for Computing Machinery, Inc. (ACM). SIGCSE's mission is to provide a forum for educators to discuss issues related to the development, implementation, and/or evaluation of computing programs, curricula, and courses, as well as syllabi, laboratories, and other elements of teaching and pedagogy.
%
\item Conference presentation at the International Conference on Physics Education (ICPE). ICPE brings together people working in physics educational research and in physics education at all types of schools from all over the world to enable them to share and exchange our experience.
%
\item Journal article in Physics Education. Physics Education is the international journal for everyone involved with the teaching of physics in schools and colleges. The articles reflect the needs and interests of secondary school teachers, teacher trainers and those involved with courses up to introductory undergraduate level.
%
\item Small workshops at the Association for Biology Laboratory Education (ABLE) conferences. ABLE's mission is to promote information exchange among university and college educators actively concerned with teaching biology in a laboratory setting. Each conference brings together a group of selected presenters with about 140 participants from university and college biology departments throughout Canada and the U.S. In three very full days, the participants are actively involved in four 3-hour ``hands on" workshops and several shorter ``mini" workshops. The workshop presenters provide all of the essential information and experiences that the potential user of the laboratory would require in order to ``take it home" and use the exercise in their own teaching program. The workshops are published in Tested Studies for Laboratory Teaching, the conference proceedings published by ABLE. Selected articles are also available online. 
%
\item Conference presentations through the Society for Industrial and Applied Mathematics (SIAM). SIAM's Education Committee holds a session at the Joint Math Meeting that focuses on educational issues. They also work to provide curriculum input for programs in mathematical sciences or computational sciences as well as at a national level, including K-12.
%http://www.siam.org/students/resources/report.php
%
\item Conference presentations through the American Nuclear Society (ANS). ANS's Education Training and Workforce Development Division hosts tracks at National Meetings that would be appropriate for this work. 
\end{itemize}

\end{document}