\documentclass[12pt]{article}

\usepackage[top=0.75in, bottom=0.75in, left=1in, right=1in]{geometry}
\pagestyle{empty}
\usepackage{tabu}

\begin{document}

\begin{center}
{NSF IWBW Summaries}
\end{center}

\setlength{\unitlength}{1in}
\begin{picture}(6,.1) 
\put(0,0) {\line(1,0){6.25}}         
\end{picture}

\renewcommand{\arraystretch}{2}
%\vspace{2em}
%\noindent \underline{Workshop goal}: enhance the participants' knowledge of essential elements of an educational proposal, or education component of more general proposal, and their understanding of strategies for developing more effective proposals. 

\noindent \textbf{Proposal Strategies}

\vspace{2em}
\noindent Competitive proposals have
\begin{itemize}
\item an understanding of the review process reflected in them
\item great idea(s) that addresses important problems in the field, are embedded in relevant theory and literature, and are significant to relevant stakeholders
\item well designed project, including management and implementation plans
\item a coherent narrative that integrates each component of the project, including evaluation
\end{itemize}

\vspace{1em}
\noindent Review criteria -- must address \textbf{both}
\begin{enumerate}
\item intellectual merit (potential to advance knowledge)
\item broader impacts (potential to benefit society and contribute to the achievement of specific, desired outcomes)
\end{enumerate}

\vspace{1em}
\noindent General:
\begin{itemize}
\item What is the potential for the activity to meet these criteria?
\item Does it explore creative, original, or potentially transformative concepts?
\item Is it well-reasoned and well-organized? Can the project assess success?
\item How well qualified is the team?
\item Are there adequate resources?
\end{itemize}

\vspace{1em}
\noindent \underline{Elements of a competitive proposal:}
\vspace{1em}

\noindent * Outline goals and objectives (what)

\vspace{1em}
%%%%%%%
\noindent * Use rationale to provide (why)
\begin{itemize}
\item background and context
  \begin{itemize}
  \item how does the proposed work fit into and relate to prior
        work of others? Of the applicant?
  \item How does the proposed work fit into and relate to relevant
        theories?
  \end{itemize}
\item justification and significance
  \begin{itemize}
  \item How does the proposal incorporate the current understanding
        of teaching and learning?
  \item '' new disciplinary knowledge?
  \item '' address an emerging area or known problem?
  \item '' address an industry/scientific/societal need?
  \item How easily is the proposed work transportable to other
        institutions and contexts?
  \item What are the potential contributions to the teaching and
        learning knowledge base?
  \item What are the potential limitations of alternate approaches?
  \end{itemize}
\item connect the goals and objectives to the project plan
\item http://ies.ed.gov/pdf/CommonGuidelines.pdf (outlines types of research and associated expected justification and evidence)
\end{itemize}
%%%%%%%
* Project Plans (how)
\begin{itemize}
\item implementation plan (how and when things are done): strategies; products; equipment and resources; timeline\\
\emph{Timeline} shows connections between the other pieces of the plan (management, eval, etc.)  
\item management plan (who is doing something when): demonstrates responsibility and appropriateness $\rightarrow$ high probability of success.
 \begin{itemize}
 \item highlight qualifications of researchers
 \item each PI/co-PI has specific tasks
 \item \textbf{independent, external evaluator is used}
 \end{itemize}
\item dissemination plan
\item evaluation plan
\end{itemize}

%--------------------------------------------------------------------------------
\vspace{2em}
\noindent \textbf{Transportability and Impact}

\vspace{1em}
\noindent This is written with the perspective of beefing up our dissemination section.

\vspace{1em}
\noindent Features that make projects transportable:
\begin{itemize}
\item built in flexibility (we cover this well)
\item identification of how the approach can be used with other teaching
  styles, other models, other courses, or other disciplines (we cover this
  well)
\item clearly state learning goals and objectives (addressed elsewhere)
\item address a common need (addressed elsewhere)
\item minimize specialized equipment needs and implementation costs (this is
  pretty easy for us to address)
\item collect convincing evaluation data (discussed elsewhere)
\item provide options for gradual scale up (natural feature of the bootcamps)
\item involve faculty at other sites (done)
\end{itemize}

\vspace{1em}
\noindent To make materials that might be adopted by other faculty, the adaptability
needs to be made clear. 
\begin{itemize}
\item materials are easily modifiable (check)
\item clearly explain how these materials are transferable (easy to do in our
  final products)
\item give suggested modifications for different situations (similar to
  above)
\item Encouraging, facilitation, and enabling others 
  \begin{itemize}
  \item materials need to be known about, available, and support for using
    them should be given. 
  \item there should be extensive reach and an ability for people to
    contribute back.
  \end{itemize}
\end{itemize}

\vspace{1em}
\noindent Things that make a project transformative (that apply to us):
\begin{itemize}
\item addresses a problem that the community judges to be significant and
  important.
\item develops effective products, curriculum materials, etc.
\item has transportable elements that other educators could adopt or adapt.
\item is relatively easy and inexpensive to implement.
\item has a plan to engage and enable the appropriate academic community.
\item makes strategic use of existing dissemination practices and introduces
  new ones.
\item has an extensive evaluation component.
\end{itemize}


%------------------------------------------------------------------------------
\vspace{2em}
\noindent \textbf{Broader Impacts}

\vspace{1em}
\noindent Categories of desired BI outcomes
\begin{itemize}
\item full participation of women, persons with disabilities, and
  under-represented minorities in STEM
\item improved STEM education and educator development at any level
\item increased public scientific literacy and engagement
\item improved well-being of individuals in society
\item development of a diverse, globally competitive STEM workforce
\item increased partnerships between academia, industry, and others
\item improved national security
\item increased economic competitiveness of the United States
\item enhanced infrastructure for research and education
\end{itemize}

\vspace{1em}
\noindent Intellectual Merit and Broader Impacts activities should be a whole and make
sense. Discussion of publications, etc.\ should also be here. Need to have an
evaluation plan associated with the BI. Outreach activities? Dissemination
needs to be active.

\vspace{1em}
\noindent Include rational about why that under-represented group was selected and talk
about how your approach makes sense (couch in research if possible). Think
through issues about access, e.g., off-hours activities can limit participation.

\vspace{1em}
\noindent Evaluation plan for broadening participation strategies is needed.

\vspace{2em}
\noindent ----- Aside from the CommonGuidelines document -----\\
It looks like we're doing

\underline{Research Type \#5}: Effectiveness Research examines effectiveness of a strategy or intervention under circumstances that would typically prevail in the target context. The importance of ``typical" circumstances means that there should not be more substantial 
developer support than in normal implementation, and there should not be substantial developer involvement in the evaluation of the strategy or intervention.

\underline{Impact Research} includes designs that eliminate or reduce bias arising from self-selection into treatment and control 
conditions, clearly specified outcome measures, adequate statistical power to detect effects, and data on implementation of the intervention or strategy and the counterfactual condition.

\end{document}
